Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid

### 2016-02-04 Appendix.tex
last minute changes
appendix update

\onecolumn
\centering
\appendix
   \subsection{Notations}

	\begin{tabular}{ | l | r | }
   						N & Number of Prosumers \\
   						$ P_i^P(t) $ & Available production of agent i at time t \\
   						$ P_i^D(t) $ & Consumption of agent i at time t \\ 
  						$ P_i(t) = P_i^P(t) - P_i^D(t) $ & Available extra-production of agent i at time t \\
  						$  WT_i $ & Wind turbine set of agent i \\
  						$ PV_i $ & Solar array set of agent i \\
  						$ DER_i = WT_i \cup PV_i $ & DER of agent i \\
  						$ \nu_i(t) $ & wind speed (in $ m s^{-2} $) at agent i's location and time t \\
  						$ \Psi_i(t) $ & solar irradiance (in $ W m^{-2} $) at agent i's location and time t \\
  						$ \mathcal{F}_{WT} $ & power curve for the wind turbines \\
  						$ \mathcal{F}_{PV} $ & power curve for the solar arrays \\
   						$ P_S^{CRCT} $ & Contract value of coalition S \\
   						$ \phi \in \left[ 0,1 \right] $ & Reliability Threshold \\
   						$ P^{MIN} $ & Minimum Contract value \\
   						$ \alpha $ & Parameter that controls the coalitions sizes \\
   						$ \epsilon $ & Correlation graph filtering threshold \\
   						$ N_{COAL} $ & Number of coalitions \\
   						$ \mathcal{R}_S $ & Resilience of coalition S \\
   						$ \lambda $ & unitary price rate for electricity \\
   						$ \delta_S(i) $ & marginal utility of agent i in coalition S \\
   						$ \rho_{ij} $ & Pearson's correlation coefficient between $ P_i$ and $P_j$ \\
   						$ \Theta_k(G) $ & set of non overlapping cliques of size k in graph G \\
   						$ \mathcal{R}_S $ & resilience of coalition S
 	\end{tabular}

\twocolumn
\subsection{Algorithms}

@@

@@

\bibliographystyle{IEEEtran}  
\bibliography{Article}

\end{document}

>>>>

%\onecolumn
\appendix

%\twocolumn
\subsection{Algorithms}

@@

 \subsection{Notations}

Figure \ref{fig:notations} presents all the notations used in the article.

\begin{figure*}
\begin{minipage}{\textwidth}
\centering
	\begin{tabular}{ | l | r | }
   						N & Number of Prosumers \\
   						$ P_i^P(t) $ & Available production of agent i at time t \\
   						$ P_i^D(t) $ & Consumption of agent i at time t \\ 
  						$ P_i(t) = P_i^P(t) - P_i^D(t) $ & Available extra-production of agent i at time t \\
  						$  WT_i $ & Wind turbine set of agent i \\
  						$ PV_i $ & Solar array set of agent i \\
  						$ DER_i = WT_i \cup PV_i $ & DER of agent i \\
  						$ \nu_i(t) $ & wind speed (in $ m s^{-2} $) at agent i's location and time t \\
  						$ \Psi_i(t) $ & solar irradiance (in $ W m^{-2} $) at agent i's location and time t \\
  						$ \mathcal{F}_{WT} $ & power curve for the wind turbines \\
  						$ \mathcal{F}_{PV} $ & power curve for the solar arrays \\
   						$ P_S^{CRCT} $ & Contract value of coalition S \\
   						$ \phi \in \left[ 0,1 \right] $ & Reliability Threshold \\
   						$ P^{MIN} $ & Minimum Contract value \\
   						$ \alpha $ & Parameter that controls the coalitions sizes \\
   						$ \epsilon $ & Correlation graph filtering threshold \\
   						$ N_{COAL} $ & Number of coalitions \\
   						$ \mathcal{R}_S $ & Resilience of coalition S \\
   						$ \lambda $ & unitary price rate for electricity \\
   						$ \delta_S(i) $ & marginal utility of agent i in coalition S \\
   						$ \rho_{ij} $ & Pearson's correlation coefficient between $ P_i$ and $P_j$ \\
   						$ \Theta_k(G) $ & set of non overlapping cliques of size k in graph G \\
   						$ \mathcal{R}_S $ & resilience of coalition S
 	\end{tabular}

\end{minipage}
\caption{Notations}
\label{fig:notations}
\end{figure*}

@@

\subsection{Utility function}
\label{appendix_utility}

We have the following utility function (see main article) that attributes a real value to any coalition S depending on the best contract value that S can propose given the grid policy $ (P^{MIN},\phi)$ :

\[ U(S) = \frac{1}{|S|^{\alpha}} \frac{ P_S^{CRCT \star} }{P^{MAX}} \]

$ U $ depends on the specific agents within the coalition S because $ P_S^{CRCT \star} $ is directly impacted by the correlations between the agents. $ U $ is also sensitive to the size of the coalition S through the term $ \frac{1}{|S|^{ \alpha } } $. The purpose of this term is to introduce in the utility function the following tradeoff : as a coalition grows, its contract value tends to grow as well but the cost of coordinating all the agents also becomes larger. The parameter $ \alpha $ is used in order to control to what extend the size of the coalition impacts its utility.

Knowing what value to use for $ \alpha $ is not trivial and is important since it influences the shape of the utility function. For some range of values, the utility might not even be a concave function of the coalition size. Given a situation with N agents and a desired number of coalitions $ N_{COAL}$, we want to obtain some estimation of $ \alpha $ such that the utility function is concave, and such that the maximum lies around a desired size $ \bar{N} = \frac{N}{N_{COAL}} $. We thus target a local maximum in this balanced region of the space by shaping the utility function through $ \alpha$.
\\
In order to obtain an analytical expression for $ \alpha $, we use two simplifying assumptions. First, we consider the distributions to be normal distributions, such that the optimal contract value can be written in the following way :

\[ P_S^{CRCT \star } = \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi) \]

Therefore, we can writte the utility function as :

\[ U(S) = \frac{1}{|S|^{\alpha} P^{MAX} } \left[ \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi) \right] \]

With :

\[ \left\{ \begin{array}{lll} \mu_{S} = \sum_{i \in S} \mu_i \\ \sigma_{S} = \sqrt{ \sum_{i \in S} \sigma_i^2 + 2 \sum_{i} \sum_{j} \rho_{ij} \sigma_i \sigma_j } \end{array} \right. \]

We then make a second assumption for simplifying this expression : we consider that :

\[ \left\{ \begin{array}{lll} \forall i \in \mathcal{A}\ \mu_i = \bar{ \mu } = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i  \\ \forall i \in \mathcal{A}\ \sigma_i = \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i \\ \forall i,j \in \mathcal{A}^2\  \rho_{ij} = \bar{ \rho } = \frac{2}{N(N-1)} \sum_{i \in \mathcal{A}} \sum_{j \in \mathcal{A}} \rho_{ij} \end{array} \right.\]

That is, any agent is replaced by a mean approximation. In these conditions, we have :

\[ U(S) = \frac{ \bar{ \mu } }{ |S|^{\alpha -1} P^{MAX} } - \frac{ \sqrt(2) \bar{ \sigma } }{ |S|^{\alpha} P^{MAX} } \sqrt{|S| + \bar{ \rho } |S| ( |S| - 1 ) } \]

$ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $. Depending on $ \alpha $ (and $ \bar{ mu },\ \bar{ \sigma },\ \bar{ rho }, and\ P^{MAX} $), the utility function can have very different shapes.
\\
We find $\alpha^{\star} $ by solving :

\[ \left\{ \begin{array}{lll} \left| \frac{ \partial U }{ \partial |S| } \right|_{|S|=\bar{N}} = 0 \\ \left| \frac{ \partial^2 U }{ \partial |S|^2 } \right|_{\alpha = \alpha^{\star}} \leq 0 \end{array} \right. \]

Solving this equation with a formal calculus software yields the equation of the paper.

\bibliographystyle{IEEEtran}  
\bibliography{Article}
\end{document}
